Data Exploration: IUCN and AquaMaps

### 2015-11-13 data_explore.Rmd
generating new plots for document: amount of overlap vs ratio of area, etc

```{r,echo=F,warning=FALSE,message=F}

@@

```{r,echo=F,warning=FALSE,message=F,results='hide'}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN species
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), col_types = 'cdddd', progress = TRUE)%>%as.data.frame()%>%filter(presence!=5)

@@

```{r,echo=F,warning=FALSE,message=F}

@@

```{r,echo=F,eval=FALSE,warning=FALSE,message=F}

spp_list = spp_list %>%
              mutate(total_area = NA,
                     am_area    = NA,
                     iucn_area  = NA,
                     perc       = NA,
                     am_perc    = NA,
                     iucn_perc  = NA,
                     sm_perc    = NA, #filled with % of smaller range that is within the larger range
                     sm_range   = NA) #what range is smaller

#there's likely a better way to do this than a for loop (lapply or sapply?)
for (i in 1:nrow(spp_list)){

  spp     <- spp_list[i, ]$sciname #species name

  spp_map <- get_spp_map(spp) #get species map

###get just the aquamaps cells
   am_spp_map = spp_map%>%
                   filter(!is.na(am_prob))%>%
                    dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
                     unique()

  if(TRUE %in% duplicated(am_spp_map$loiczid)){
                  am_spp_map = am_spp_map%>%
                                dplyr::select(loiczid,am_sid)%>%
                                  unique()}

@@

  spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
              cellStats(.,stat='sum')

###  get the iucn cells
  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
        unique()

@@

  spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
                cellStats(.,stat='sum')

@@

  spp_list[i,]$total_area = area(fullRange,na.rm=T)%>%
                              cellStats(.,stat='sum')

@@

   #iucn map cells
  cells_iucn<-iucn_map%>%
              filter(iucn_pres==TRUE)%>%
              .$loiczid
  cells_am <- am_spp_map%>%
                filter(!is.na(am_sid))%>%
                .$loiczid

  cells_total = unique(spp_map$loiczid)
  cells_overlap = intersect(cells_iucn,cells_am)

  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
#    r_iucn_spp <- subs(loiczid_raster, 
#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
#                      by = 'loiczid', 
#                       which = 'iucn_pres', 
#                       subsWithNA = TRUE)
#   
  #percent overlap
  perc = (length(cells_overlap)/length(cells_total))*100
  spp_list[i,]$perc = perc

  am_only = setdiff(cells_am,cells_iucn)
  iucn_only = setdiff(cells_iucn,cells_am)

  perc_am = (length(am_only)/length(cells_total))*100
  perc_iucn = (length(iucn_only)/length(cells_total))*100

  spp_list[i,]$am_perc = perc_am
  spp_list[i,]$iucn_perc = perc_iucn

  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n',i,spp,perc,perc_am,perc_iucn))

@@

#remove non IUCN species and log total area (for use later on)
spp_list = spp_list%>%
            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids')%>%
             mutate(log_total_area = log(total_area))

write.csv(spp_list,file='spp_list_w_area_percent_noextinct.csv')
```

@@

```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}
# look at number of species per group

spp_list = read.csv('spp_list_w_area_percent_noextinct.csv')

df = spp_list%>%
      mutate(diff = iucn_area - am_area)

df <- within(df,spp_group <- factor(spp_group,levels=names(sort(table(spp_group),increasing=TRUE))))

g<-ggplot(df,aes(x=spp_group))+
  geom_bar(stat='bin')+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+ coord_flip()+
  labs(x='Species Group', y='Number of Species',title='Number of species in both AquaMaps and IUCN datasets')

ggsave('figures/sp_groups_barchart.png',plot=g,width=8, height=4, dpi=100)
print(g)

@@

```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}

dot = ggplot(df,aes(iucn_area,am_area,color=iucn_category))+ geom_point() + geom_abline()+
      labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png',plot=dot,width=8, height=4, dpi=100)
print(dot)
```

There are large differences so to get a better sense, here is the same comparison but with area log transformed.
````{r echo=F,warning=F,message=F,warning=FALSE,message=F}

df = df%>%
      mutate(am_log = log(am_area),
             iucn_log = log(iucn_area))

all_status <- ggplot(df,aes(iucn_log,am_log,color=iucn_category)) + 
  geom_point() + geom_abline() +
  labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
        title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')
ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png',plot=all_status,width=8,height=4,dpi=100)
print(all_status)
```

Data split into species groups and by IUCN category
```{r,echo=F,warning=F,message=F}

groups <- ggplot(df, aes(iucn_log, am_log,color=iucn_category)) + geom_point() +facet_wrap(~spp_group) + 
        geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
                             title='Comparing differences in area across species groups', fill='IUCN Category')

ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png',plot=groups,width=8,height=6,dpi=100)
print(groups)
```
***

@@

```{r,echo=F,warning=FALSE,message=F}

spp = 'Chromis alpha'
map <- get_spp_map(spp)

 am_map <- map%>%
            filter(!is.na(am_prob))

@@

    r_am_spp[!is.na(r_am_spp)]<-1

@@

#plot
    plot(r_iucn_spp,col='mediumpurple',useRaster=FALSE,axes=FALSE,box=FALSE,legend=FALSE)
  plot(r_am_spp,col='orange',main="Yellow-speckled Chromis", useRaster=FALSE,add=T,axes=FALSE,box=FALSE,legend=FALSE)
  map('world', col = 'gray95', fill = T, border = 'gray80', add = TRUE)
  legend("topright",legend=c("AquaMaps","IUCN"),fill=c('orange','mediumpurple'))

@@

```{r,echo=F,warning=FALSE,message=F}

  df <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
          dplyr::select(sciname, spp_group, perc.type, percent) %>%
            unique()

  perc_c = df %>%
            filter(sciname == spp)

   p1 = perc_c %>% dplyr::select(percent)

@@

For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2,],2)`% is covered just by AquaMaps and `r round(p1[3,], 2)`% is covered just by IUCN.

@@

```{r,echo=F,warning=FALSE,message=F}

#look at density plot

overlap = df%>%dplyr::filter(perc.type=='perc')

all<-ggplot(overlap,aes(percent))+
    geom_histogram(aes(y=..count../1926),binwidth=5,alpha=.8)+
    labs(x='% total range that is shared b/w both datasets', y='Proportion of Total Species')
ggsave('figures/Overlap-PercTotalRange.png')
print(all)
```

Split up by species - not sure if this is useful or not
```{r,echo=F, message = F, warning = F}

sp <- ggplot(overlap,aes(percent))+geom_histogram(binwidth=5,alpha=.8) +
  facet_wrap(~spp_group) +
  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')+ 
       theme(strip.text.x = element_text(size = 8))

ggsave('figures/Overlap-PercTotalRange_spGroups.png')
print(sp)

@@

```{r pct small within large, echo=F, warning= F}

d <- spp_list %>%
  dplyr::select(sciname, sm_perc, sm_range) %>%
      unique()

sm_plot <- ggplot(d, aes(sm_perc, fill = sm_range)) + 
       geom_histogram(aes(y=..count../sum(..count..)),binwidth = 4, alpha = .8) +
        scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
        labs(x = 'Percent of smaller range that falls within the larger range', y = 'Proportion of all species (n=1962)')
ggsave('figures/PercSmallRangeInLarger.png',width=6,height=4)

@@

Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
``` {r pct overlap vs area, echo = FALSE, warning = F, message=F}

reg <- ggplot(spp_list, aes(sm_perc,total_area, color = sm_range)) + 
  geom_point() + 
  geom_abline() + 
  labs(y = 'Total Area (km2)', x = 'Percent overlap smaller range in larger', 
       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')

ggsave('figures/AreavsPercOverlap.png',width=6,height=4)

@@

``` {r, echo = FALSE, warning = F, message=F}

@@

summary(lm(total_area~sm_perc,spp_list))
```

@@

Perent coverage as function of area, clipping to area below 100,000,000 km^2:
``` {r, echo = F, warning = F, message = F}
# remove all area greater than 100000000

@@

ggsave('figures/AreavsPercOverlap_2.png',width=6,height=4)

@@

Percent coverage as a function of log(area)
``` {r, echo = F, warning = F, message = F}
# Log transform area

@@

ggsave('figures/AreavsPercOverlap_log.png',width=6,height=4)

@@

``` {r examine regressions, echo = FALSE, warning = F}

@@

###(5) Examining data broken into quartiles to see if smaller patterns emerge
``` {r examine quartiles, echo = FALSE, warning = F}

@@

qtiles <- c(quantile(spp_list$total_area, 0.0,na.rm=T),
            quantile(spp_list$total_area, .25,na.rm=T),
            quantile(spp_list$total_area, .50,na.rm=T),
            quantile(spp_list$total_area, .75,na.rm=T),
            quantile(spp_list$total_area, 1.0,na.rm=T))

@@

  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
  geom_point() + 
  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
       title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
ggsave(paste0('figures/AreavsOverlap_quantile_',i,'.png'),width=6,height=4)
print(qtile_plot)
}
# First quartile regression:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  1423917     101053  14.091   <2e-16 ***
#     sm_perc         1188       1555   0.764    0.445    
#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
#
# second quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
#
# third quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
#     sm_perc        64710       8159   7.931  1.5e-14 ***
#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
#
# fourth quartile:
#                   Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
#     sm_perc        1911286     215216   8.881  < 2e-16 ***
#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381

@@

```{r aquamaps probabilities, warning=F,message=F,cache.lazy=FALSE,cache=T,verbose=F,results=F,echo=F,include=F}

@@

dir_N = c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
          'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_aquamaps = file.path(dir_N,"git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")

spp_cells = fread(input=file.path(dir_aquamaps,"tables/ohi_hcaf_species_native.csv"),verbose=F,header=T,
                  colClasses=c('NULL','character',NA,'numeric','NULL','NULL')) #1 min 49 sec

setnames(spp_cells,c('SpeciesID','CsquareCode','probability'),c("species_id","csquare_code","probability"))

spp_cells = as.data.frame(spp_cells)

spp       = fread(input=file.path(dir_aquamaps,"tables/ohi_speciesoccursum.csv"),verbose=F,header=T)%>%
              mutate(scientific = paste(Genus,Species,sep=" "))
spp = as.data.frame(spp)

@@

```{r,warning=F,message=F, echo=F}

#create dataframe that gives me all species, their total number of cells, total number of cells >=0.4 prob, differnece between the two and threat
df = spp_cells%>%
      group_by(species_id)%>%
        summarize(cells_0.4 = sum(probability>=0.4),
                  cells_t = sum(probability>=0))%>%
          mutate(prop = cells_0.4/cells_t)%>%
            left_join(spp,by=c('species_id'='SPECIESID'))%>%
          select(scientific,species_id,cells_0.4,cells_t,prop,iucn_code)

hist <- ggplot(df,aes(prop,fill=iucn_code))+geom_histogram(aes(y=..count../sum(..count..)),binwidth = 0.05, alpha = .7)+labs(x='Proportion of range area included when selecting a threshold of 0.4',y='Proportion of all aquamaps species (17,348 total)')
ggsave('figures/AMThreshold.png',width=6,height=4)
print(hist)

>>>>

```{r, echo = FALSE, warning = FALSE, message = FALSE}

@@

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN species
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
                           col_types = 'cdddd', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5)

@@

```{r species_map_function, echo = FALSE, warning = FALSE, message = FALSE}

@@

```{r create_species_list_w_area, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}

spp_list <- spp_list %>%
              mutate(total_area = NA,
                     am_area    = NA,
                     iucn_area  = NA,
                     perc       = NA,
                     am_perc    = NA,
                     iucn_perc  = NA,
                     sm_perc    = NA, #filled with % of smaller range that is within the larger range
                     sm_range   = NA) #what range is smaller

#there's likely a better way to do this than a for loop (lapply or sapply?)
for (i in 1:nrow(spp_list)){
  spp     <- spp_list[i, ]$sciname #species name
  spp_map <- get_spp_map(spp) #get species map

###get just the aquamaps cells
   am_spp_map <- spp_map%>%
                   filter(!is.na(am_prob))%>%
                    dplyr::select(loiczid, am_sid, am_prob)%>% #add in am_prob when we want to look at probability
                     unique()

  if(any(duplicated(am_spp_map$loiczid))){
    am_spp_map <- am_spp_map %>%
      dplyr::select(loiczid, am_sid) %>% 
      unique()
    }

@@

  spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
              cellStats(., stat='sum')

###  get the iucn cells
  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
        unique()

@@

  spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
                cellStats(., stat='sum')

@@

  spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE)%>%
                              cellStats(., stat='sum')

@@

   #iucn map cells
  cells_iucn<-iucn_map%>%
              filter(iucn_pres == TRUE)%>%
              .$loiczid
  cells_am <- am_spp_map%>%
                filter(!is.na(am_sid))%>%
                .$loiczid

  cells_total <- unique(spp_map$loiczid)
  cells_overlap <- intersect(cells_iucn, cells_am)

  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
#    r_iucn_spp <- subs(loiczid_raster, 
#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
#                      by = 'loiczid', 
#                       which = 'iucn_pres', 
#                       subsWithNA = TRUE)
#   
  #percent overlap
  perc <- (length(cells_overlap) / length(cells_total)) * 100
  spp_list[i, ]$perc <- perc

  am_only <- setdiff(cells_am, cells_iucn)
  iucn_only <- setdiff(cells_iucn, cells_am)

  perc_am <- (length(am_only) / length(cells_total)) * 100
  perc_iucn <- (length(iucn_only) / length(cells_total)) * 100

  spp_list[i, ]$am_perc <- perc_am
  spp_list[i, ]$iucn_perc <- perc_iucn

  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') #what dataset has the smaller range
  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 #this is the total percent of smaller range encompassed in the larger range
cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', i, spp, perc, perc_am, perc_iucn))

@@

#remove non IUCN species and log total area (for use later on)
spp_list <- spp_list%>%
            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids')%>%
             mutate(log_total_area = log(total_area))

write.csv(spp_list, file='spp_list_w_area_percent_noextinct.csv')
```

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# look at number of species per group

spp_list <- read.csv('spp_list_w_area_percent_noextinct.csv')

spp_df <- spp_list %>%
  mutate(diff = iucn_area - am_area)

spp_df <- within(spp_df, spp_group <- factor(spp_group, levels = names(sort(table(spp_group), increasing = TRUE))))

barchart_sp_groups <- ggplot(spp_df, aes(x = spp_group)) +
  geom_bar(stat = 'bin') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
  coord_flip() +
  labs(x = 'Species Group', 
       y = 'Number of Species', 
       title = 'Number of species in both AquaMaps and IUCN datasets')

ggsave('figures/sp_groups_barchart.png', 
       plot = barchart_sp_groups, 
       width = 8, height = 4, dpi = 100)

print(barchart_sp_groups)

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE, warning = FALSE, message = FALSE}

scatter_am_iucn_area <- ggplot(spp_df, aes(iucn_area, am_area, color = iucn_category)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = 'Area (km2) from IUCN dataset', 
       y = 'Area (km2) from AquaMaps dataset', 
       title = 'Comparing total area of species range maps')

ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png', 
       plot = scatter_am_iucn_area, 
       width = 8, height = 4, dpi = 100)

print(scatter_am_iucn_area)
```

There are large differences so to get a better sense, here is the same comparison but with area log transformed.
``` {r echo = FALSE, warning = FALSE, message = FALSE}

spp_df <- spp_df %>%
  mutate(am_log = log(am_area),
         iucn_log = log(iucn_area))

scatter_am_iucn_area_log <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = 'IUCN Area (log transformed)', 
       y = 'AquaMaps Area (log transformed)',
       title = 'Comparing area of range maps\n (all area values are km2 log transformed)', 
       fill = 'IUCN Category')

ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png', 
       plot = scatter_am_iucn_area_log, 
       width = 8, height = 4, dpi = 100)

print(scatter_am_iucn_area_log)
```

Data split into species groups and by IUCN category
```{r, echo = FALSE, warning = FALSE, message = FALSE}

scatter_am_iucn_alog_gp <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
  geom_point() +
  facet_wrap(~spp_group) + 
  geom_abline() + 
  labs(x = 'IUCN Area (log transformed)', 
       y = 'AquaMaps Area (log transformed)',
       title = 'Comparing differences in area across species groups', 
       fill = 'IUCN Category')

ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png', 
       plot = scatter_am_iucn_alog_gp, 
       width = 8, height = 6, dpi = 100)

print(scatter_am_iucn_alog_gp)
```
***

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE}

spp <- 'Chromis alpha'
map <- get_spp_map(spp)

am_map <- map %>%
  filter(!is.na(am_prob))

@@

r_am_spp[!is.na(r_am_spp)] <- 1

@@

### plot raster of species: IUCN map, AM map, world map, legend
plot(r_iucn_spp, col = 'mediumpurple', 
     useRaster = FALSE, 
     axes = FALSE, 
     box = FALSE, 
     legend = FALSE)

plot(r_am_spp, col = 'orange', 
     main = "Yellow-speckled Chromis", 
     useRaster = FALSE, 
     add = TRUE, 
     axes = FALSE, 
     box = FALSE, 
     legend = FALSE)

map('world', col = 'gray95', 
    fill = TRUE, 
    border = 'gray80', 
    add = TRUE)

legend("topright", 
       legend = c("AquaMaps", "IUCN"), 
       fill = c('orange', 'mediumpurple'))

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE}

  spp_df2 <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
    dplyr::select(sciname, spp_group, perc.type, percent) %>%
    unique()

  perc_c <- spp_df2 %>%
    filter(sciname == spp)

  p1 <- perc_c %>% 
    dplyr::select(percent)

@@

For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2, ], 2)`% is covered just by AquaMaps and `r round(p1[3, ], 2)`% is covered just by IUCN.

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE}

### look at density plot

overlap <- spp_df2 %>%
  dplyr::filter(perc.type == 'perc')

hist_overlap_pct <- ggplot(overlap, aes(percent)) +
  geom_histogram(aes(y = ..count../1926), binwidth = 5, alpha = .8) +
  labs(x = '% total range that is shared b/w both datasets', 
       y = 'Proportion of Total Species')

ggsave('figures/Overlap-PercTotalRange.png')

print(hist_overlap_pct)
```

Split up by species - not sure if this is useful or not
```{r, echo = FALSE, message = FALSE, warning = FALSE}

hist_overlap_pct_gp <- ggplot(overlap, aes(percent)) + 
  geom_histogram(binwidth = 5, alpha = .8) +
  facet_wrap(~spp_group) +
  labs(x = 'Percent of total range that is shared between both datasets', 
       y = 'Number of Total Species') + 
  theme(strip.text.x = element_text(size = 8))

ggsave('figures/Overlap-PercTotalRange_spGroups.png')
print(hist_overlap_pct_gp)

@@

```{r pct small within large, echo = FALSE, warning = FALSE}

spp_df4 <- spp_list %>%
  dplyr::select(sciname, sm_perc, sm_range) %>%
  unique()

sm_plot <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
  scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
  labs(x = 'Percent of smaller range that falls within the larger range', 
       y = 'Proportion of all species (n = 1962)')

ggsave('figures/PercSmallRangeInLarger.png', width = 6, height = 4)

@@

Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}

reg <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
  geom_point() + 
  geom_abline() + 
  labs(y = 'Total Area (km2)', 
       x = 'Percent overlap smaller range in larger', 
       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')

ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)

@@

``` {r, echo = FALSE, warning = FALSE, message = FALSE}

@@

summary(lm(total_area~sm_perc, spp_list))
```

@@

Perent coverage as function of area, clipping to area below 100,000,000 km^2:
``` {r, echo = FALSE, warning = FALSE, message = FALSE}
# remove all area greater than 100000000

@@

ggsave('figures/AreavsPercOverlap_2.png', width = 6, height = 4)

@@

Percent coverage as a function of log(area)
``` {r, echo = FALSE, warning = FALSE, message = FALSE}
# Log transform area

@@

ggsave('figures/AreavsPercOverlap_log.png', width=6, height=4)

@@

``` {r examine regressions, echo = FALSE, warning = FALSE}

@@

###(5) Examining data broken into quartiles to see if smaller patterns emerge
``` {r examine quartiles, echo = FALSE, warning = FALSE}

@@

qtiles <- c(quantile(spp_list$total_area, 0.0, na.rm = TRUE),
            quantile(spp_list$total_area, .25, na.rm = TRUE),
            quantile(spp_list$total_area, .50, na.rm = TRUE),
            quantile(spp_list$total_area, .75, na.rm = TRUE),
            quantile(spp_list$total_area, 1.0, na.rm = TRUE))

@@

  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
    geom_point() + 
    labs(x = 'Total Area (km2)', 
         y = 'Percent overlap smaller range in larger', 
         title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))

  ggsave(paste0('figures/AreavsOverlap_quantile_', i, '.png'), width = 6, height = 4)
  print(qtile_plot)
}
# First quartile regression:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  1423917     101053  14.091   <2e-16 ***
#     sm_perc         1188       1555   0.764    0.445    
#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
#
# second quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
#
# third quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
#     sm_perc        64710       8159   7.931  1.5e-14 ***
#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
#
# fourth quartile:
#                   Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
#     sm_perc        1911286     215216   8.881  < 2e-16 ***
#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381

@@

```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}

@@

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_aquamaps <- file.path(dir_N, "git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")

spp_cells <- fread(input = file.path(dir_aquamaps, "tables/ohi_hcaf_species_native.csv"), verbose = FALSE, header = TRUE, 
                  colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec

setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c("species_id", "csquare_code", "probability"))

spp_cells <- as.data.frame(spp_cells)

spp       <- fread(input = file.path(dir_aquamaps, "tables/ohi_speciesoccursum.csv"), 
                  verbose = FALSE, header = TRUE) %>%
              mutate(scientific = paste(Genus, Species, sep = " "))
spp <- as.data.frame(spp)

@@

```{r, warning = FALSE, message = FALSE, echo = FALSE}

#create dataframe that gives me all species, their total number of cells, total number of cells >= 0.4 prob, differnece between the two and threat
spp_df3 <- spp_cells %>%
  group_by(species_id) %>%
  summarize(cells_0.4 = sum(probability >= 0.4),
            cells_t = sum(probability >= 0)) %>%
  mutate(prop = cells_0.4 / cells_t) %>%
  left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
  select(scientific, species_id, cells_0.4, cells_t, prop, iucn_code)

hist <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', y = 'Proportion of all aquamaps species (17,348 total)')

ggsave('figures/AMThreshold.png', width = 6, height = 4)

print(hist)
